### 2023-07-17 paper/sec-dis-pseudodata.tex
sect 2

\section{DIS projections for LHC far-forward experiments}
\label{sec:dis_pseudodata}

Here we describe the procedure adopted in order 
to generate projections for the kinematic coverage
and experimental uncertainties associated to  measurements
of neutrino-nucleus scattering at the LHC far-forward experiments.
%
First, we summarise the theoretical description of differential
neutrino scattering in terms of DIS structure functions and highlight
their PDF sensitivity.
%
Second, we present an  overview of the operative and proposed
LHC far-forward neutrino detectors that
will be considered in the present study and indicate their
main acceptance and performance parameters.
%
By convoluting the expected muon neutrino fluxes
with the acceptance and scattering rates of each
of these detectors,
we evaluate the foreseen event yields in bins of $x$, $Q^2$,
and $E_\nu$ and therefore the associated statistical uncertainties.
%
We also estimate how the systematic uncertainties in the measurement
of final-state variables affects the experimental covariance matrix.

@@

 The calculation of differential neutrino scattering event rates
 at the LHC far-forward detectors involves two main ingredients: the energy
 and flavour dependence of the incoming neutrino flux crossing
 the detector fiducial volume, on the one hand,
 and the scattering rates within the same fiducial volume, on the other hand.
 %
 In this section, we summarise the main features of each of the existing and future
 far-forward detectors considered in this study, in particular concerning
 their acceptance, performance, and detection method.
 %
 We focus on  muon
 neutrino scattering, which benefits from the highest rates and is less
 affected by uncertainties affecting $D$-meson production.

 The kinematics of a charged-current neutrino DIS event $(x,Q^2, E_\nu)$,
 or alternatively $(x,Q^2, y)$, are uniquely specified by the measurement of three independent
 final-state variables,
 such as $\lp E_\ell,\theta_\ell, W^2\rp$ or $\lp E_\ell,\theta_\ell, E_h \rp$,
 with $E_\ell$ and $\theta_\ell$ being the energy and polar angle of the outgoing
 charged lepton and $E_h$ the total energy of the hadronic final state.
 %
 Most neutrino detectors can only access $E_h$, given that $W^2$ requires
 fully reconstructing this final state.
 %
 A measurement of  $\lp E_\ell,\theta_\ell, E_h \rp$ then fixed the DIS kinematics as:
 \bea
 E_\nu &=& E_h + E_\ell \, , \nonumber \\
 Q^2 &=& 4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp \, ,  \label{eq:dis_kinematic_mapping}\\
 x&=& \frac{4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp}{2m_N E_h} \, .\nonumber
 \eea
 These relations also indicate how systematic uncertainties affecting the measurement
 of $\lp E_\ell,\theta_\ell, E_h \rp$ translate into uncertainties in the
 reconstructed values of  $(x,Q^2, E_\nu)$, which in turn modify the expected
 event rates in a given bin.

 Table~\ref{tab:FPF_experiments} indicates,
for each of the far-forward LHC neutrino experiments considered
  in this work,  their neutrino pseudo-rapidity coverage, target material, whether
  they can identify the sign of the outgoing charged lepton,
  the acceptance for the charged lepton and hadronic final state,
  and the expected reconstruction performance.
  %
  We consider separately acceptance and performance for electron and muon
  neutrinos, while tau neutrinos have lower production rates and are neglected in this study.
  %
  For these projections we assume that FASER$\nu$ and SND@LHC acquire data
  for the full Run III period, while FASER$\nu$2, AdvSND, and FLArE take data
  for the complete HL-LHC period.
  %
  An estimate of the expected 
  systematic uncertainties is not available for FASER$\nu$ and SND@LHC, experiments
  which are nevertheless limited by statistics.
  %
  In the following we provide more details about the information collected in
  Table~\ref{tab:FPF_experiments}.

@@

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{eq:fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
We take the uncertainty to be the absolute difference betweeen the mean number of events in the bin for the smeared samples samples, ${\overline{N}}^{(i)}~$ and$~ N_{\rm ev}^{(i)}$ as calculated from Eq. \ref{MCintegration}, {\it{i.e.}}  $\delta N^{(i)} = |{\overline{N}}^{(i)} -N_{\rm ev}^{(i)}|$ .
%
Furthermore, 
since the systematic errors are treated as uncorrelated among them,
by producing samples where only one source of error is varied at a time
we can determine the systematic errors Eq.~(\ref{eq:event_yields_systematic_error}) in each bin.

>>>>

\section{Deep-inelastic scattering with LHC neutrinos}
\label{sec:dis_pseudodata}

Here we describe the procedure adopted in order 
to generate projections for the kinematic coverage
and uncertainties associated to  measurements
of neutrino-nucleus scattering at the LHC far-forward experiments.
%
First, we summarise the theoretical description of differential
neutrino scattering in terms of DIS structure functions and highlight
their PDF sensitivity.
%
Second, we present an  overview of the operative and proposed
LHC far-forward neutrino detectors that
will be considered in the present study and indicate their
acceptance and performance parameters.
%
By convoluting the expected electron and muon neutrino fluxes
with the acceptance and scattering rates of each
of these detectors,
we evaluate the event yields in bins of $x$, $Q^2$,
and $E_\nu$ and the associated statistical and systematic uncertainties.
%
Finally, we discuss the procedure adopted to generate pseudo-data
for LHC neutrino DIS and to their impact
 into proton and nuclear PDF determinations.

@@

 The calculation of differential neutrino scattering event rates
 at the LHC far-forward detectors involves two main ingredients: the energy
 and flavour dependence of the incoming neutrino flux crossing
 the detector fiducial volume, on the one hand,
 and the scattering rates within the same fiducial volume, on the other hand.
 %
 Here we summarise the main features of each of the existing and future
 far-forward detectors considered, in particular concerning
 their acceptance and performance.
 %
 We focus on  muon neutrino scattering, which benefits from the highest rates and is less
 affected by theoretical uncertainties in the production mechanism.

 The kinematics of a charged-current neutrino DIS event $(x,Q^2, E_\nu)$,
 or alternatively $(x,Q^2, y)$, are uniquely specified by the measurement of three independent
 final-state variables,
 such as $\lp E_\ell,\theta_\ell, W^2\rp$ or $\lp E_\ell,\theta_\ell, E_h \rp$,
 with $E_\ell$ and $\theta_\ell$ being the energy and polar angle of the outgoing
 charged lepton and $E_h$ the total energy of the hadronic final state.
 %
 Most neutrino detectors can only access $E_h$, given that $W^2$ requires
 fully reconstructing this final state.
 %
 A measurement of  $\lp E_\ell,\theta_\ell, E_h \rp$ then fixes the DIS kinematics as:
 \bea
 E_\nu &=& E_h + E_\ell \, , \nonumber \\
 Q^2 &=& 4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp \, ,  \label{eq:dis_kinematic_mapping}\\
 x&=& \frac{4 ( E_h + E_\ell) E_\ell \sin^2 \lp \theta_\ell/2\rp}{2m_N E_h} \, .\nonumber
 \eea
 These relations also indicate how systematic uncertainties affecting the measurement
 of $\lp E_\ell,\theta_\ell, E_h \rp$ translate into uncertainties in the
 reconstructed values of  $(x,Q^2, E_\nu)$ modifying the expected
 binned event rates.

 Table~\ref{tab:FPF_experiments} summarises,
for each of the far-forward LHC neutrino experiments considered
  in this work,  their neutrino pseudo-rapidity coverage, target material, whether
  they can identify the sign of the outgoing charged lepton,
  the acceptance for the charged lepton and hadronic final state,
  and the expected reconstruction performance.
  %
  We consider separately acceptance and performance for electron and muon
  neutrinos, while tau neutrinos have lower production rates and are neglected in this study.
  %
  For these projections we assume that FASER$\nu$ and SND@LHC acquire data
  for the full Run III period, while FASER$\nu$2, AdvSND, and FLArE take data
  for the complete HL-LHC period.
  %
  An estimate of the expected 
  systematic uncertainties is not available for FASER$\nu$ and SND@LHC, experiments
  which are nevertheless limited by statistics.
  %
  In the following we provide more details about the information collected in
  Table~\ref{tab:FPF_experiments}.

@@

We then produce a second Monte Carlo sample $\mathcal{D}_1$ starting from the events
of $\mathcal{D}_0$ which are then smeared with Gaussian distributions whose variances
are given by Eq.~(\ref{eq:fasernu2systematic_errors}).
%
The bin assignment of the events in the smeared sample $\mathcal{D}_1$ will
in general be different from those of the baseline sample.
%
The procedure is repeated $M$ times, leading to
$\mathcal{D}_k$ (with $k=1,\ldots,M$) smeared
samples each one leading to a different binned event yields
$ N_{\rm ev}^{(i)(k)}$.
%
We take the uncertainty to be the absolute difference between the mean number of events in the bin for the smeared samples samples, ${\overline{N}}^{(i)}~$ and$~ N_{\rm ev}^{(i)}$ as calculated from Eq. \ref{MCintegration}, {\it{i.e.}}  $\delta N^{(i)} = |{\overline{N}}^{(i)} -N_{\rm ev}^{(i)}|$ .
%
Furthermore, 
since the systematic errors are treated as uncorrelated among them,
by producing samples where only one source of error is varied at a time
we can determine the systematic errors Eq.~(\ref{eq:event_yields_systematic_error}) in each bin.

### 2023-07-17 paper/tables/detectors.tex
sect 2

  %-----------------------------------------------------------------
\begin{table}[t]
  \centering
  \small
  \renewcommand{\arraystretch}{1.50}
\begin{tabularx}{\textwidth}{Xccccc}
\toprule
Detector &  Rapidity &  Target & Charged Lepton ID & Acceptance  & Performance \\
\midrule
\midrule
\multirow{3}{*}{FASER$\nu$}  &  \multirow{3}{*}{ $\eta_\nu \ge 8.5$}  &   \multirow{2}{*}{Tungsten}  & \multirow{3}{*}{muons}      &   $E_\ell \gsim 100$ GeV   &      $\delta E_\ell \sim 30\% $    \\
&   &   \multirow{2}{*}{(1.1 ton)}  &       &  $\tan \theta_\ell \lsim 0.025$   &
$\delta \theta_\ell \sim 0.06$ mrad        \\
&   &     &       &  reconstructed $E_h$ \& charm ID   &      $\delta E_h \sim 30\%$     \\
\midrule
\multirow{2}{*}{SND@LHC}  & \multirow{2}{*}{ $7.2 \le \eta_\nu \le 8.4$}   &  Tungsten   &   \multirow{2}{*}{n/a}    &  $E_\mu \gsim 20 $ GeV     &    \multirow{2}{*}{n/a}    \\
  &    &  (0.83 ton)   &  &  $\theta_\mu \lsim 0.15, \theta_e \lsim 0.5$         &       \\
\midrule
\midrule
\multirow{3}{*}{FASER$\nu$2}  & \multirow{3}{*}{ $\eta_\nu \ge 8.5$}  & \multirow{2}{*}{Tungsten}    &   \multirow{3}{*}{muons}     &   $E_\ell \gsim 100$ GeV  &    $\delta E_\ell \sim 30\% $     \\
  &   &  \multirow{2}{*}{(20 ton)}   &       &  $\tan \theta_\ell \lsim 0.05$   &   $\delta \theta_\ell \sim 0.06$ mrad      \\
  &   &     &       &  reconstructed $E_h$ \& charm ID   &  $\delta E_h \sim 30\%$        \\
\midrule
\multirow{3}{*}{AdvSND-far}  &   \multirow{3}{*}{ $7.2 \le \eta_\nu \le 8.4$}  &
\multirow{2}{*}{Tungsten}   &   \multirow{3}{*}{muons}    &  $E_\mu \gsim 20 $ GeV  & \multirow{3}{*}{n/a}          \\
  &   &   \multirow{2}{*}{(5 ton)}  &        & $\theta_\mu \lsim 0.15, \theta_e \lsim 0.5$     &           \\
  &   &     &       &  reconstructed $E_h$   &           \\
\midrule
\multirow{3}{*}{FLArE ({\bf *})}  & \multirow{3}{*}{$\eta_\nu \ge 7.5$} & \multirow{2}{*}{LAr}  & \multirow{3}{*}{muons}  &  $E_\mu \gsim 2$ GeV, $E_e \lsim 2$ TeV    &    $\delta E_e \sim 5\% $ \\
&   &  \multirow{2}{*}{(10 ton)}   &   & $\theta_\mu \lsim 0.025$, $\theta_e \lsim 0.5$ &    $\delta \theta_e \sim 15 $ mrad   \\
 &   &     &  & reconstructed $E_h$  &    $\delta E_h \sim 30\% $   \\
  \bottomrule
\end{tabularx}
\vspace{0.2cm}
\caption{\small For each of the far-forward LHC neutrino experiments considered
  in this work, we indicate their neutrino pseudo-rapidity coverage, target material, whether
  they can identify the sign of the outgoing charged lepton,
  the acceptance for the charged lepton and hadronic final state,
  and the expected reconstruction performance.
  %
  We consider separately acceptance and performance for electron and muon
  neutrinos, while tau neutrinos have lower production rates and are not considered in this study.
  %
  For the FLArE detector, we assume that muons would be measured in the FASER2 spectrometer
  situated downstream in the FPF cavern.
  %
  See the description of each experiment in the text for more details.
  %
  For the projections in this work, we assume that FASER$\nu$ and SND@LHC acquire data
  for the full Run III period, while FASER$\nu$2, AdvSND, and FLArE take data
  for the complete HL-LHC period.
  \label{tab:FPF_experiments}
}
\end{table}
%-----------------------------------------------------------------

>>>>

  %-----------------------------------------------------------------
\begin{table}[t]
  \centering
  \small
  \renewcommand{\arraystretch}{1.50}
\begin{tabularx}{\textwidth}{Xccccc}
\toprule
Detector &  Rapidity &  Target & Charged Lepton ID & Acceptance  & Performance \\
\midrule
\midrule
\multirow{3}{*}{FASER$\nu$}  &  \multirow{3}{*}{ $\eta_\nu \ge 8.5$}  &   \multirow{2}{*}{Tungsten}  & \multirow{3}{*}{muons}      &   $E_\ell \gsim 100$ GeV   &      $\delta E_\ell \sim 30\% $    \\
&   &   \multirow{2}{*}{(1.1 ton)}  &       &  $\tan \theta_\ell \lsim 0.025$   &
$\delta \theta_\ell \sim 0.06$ mrad        \\
&   &     &       &  reconstructed $E_h$ \& charm ID   &      $\delta E_h \sim 30\%$     \\
\midrule
\multirow{2}{*}{SND@LHC}  & \multirow{2}{*}{ $7.2 \le \eta_\nu \le 8.4$}   &  Tungsten   &   \multirow{2}{*}{n/a}    &  $E_\mu \gsim 20 $ GeV     &    \multirow{2}{*}{n/a}    \\
  &    &  (0.83 ton)   &  &  $\theta_\mu \lsim 0.15, \theta_e \lsim 0.5$         &       \\
\midrule
\midrule
\multirow{3}{*}{FASER$\nu$2}  & \multirow{3}{*}{ $\eta_\nu \ge 8.5$}  & \multirow{2}{*}{Tungsten}    &   \multirow{3}{*}{muons}     &   $E_\ell \gsim 100$ GeV  &    $\delta E_\ell \sim 30\% $     \\
  &   &  \multirow{2}{*}{(20 ton)}   &       &  $\tan \theta_\ell \lsim 0.05$   &   $\delta \theta_\ell \sim 0.06$ mrad      \\
  &   &     &       &  reconstructed $E_h$ \& charm ID   &  $\delta E_h \sim 30\%$        \\
\midrule
\multirow{3}{*}{AdvSND-far}  &   \multirow{3}{*}{ $7.2 \le \eta_\nu \le 8.4$}  &
\multirow{2}{*}{Tungsten}   &   \multirow{3}{*}{muons}    &  $E_\mu \gsim 20 $ GeV  & \multirow{3}{*}{n/a}          \\
  &   &   \multirow{2}{*}{(5 ton)}  &        & $\theta_\mu \lsim 0.15, \theta_e \lsim 0.5$     &           \\
  &   &     &       &  reconstructed $E_h$   &           \\
\midrule
\multirow{3}{*}{FLArE ({\bf *})}  & \multirow{3}{*}{$\eta_\nu \ge 7.5$} & \multirow{2}{*}{LAr}  & \multirow{3}{*}{muons}  &  $E_\mu \gsim 2$ GeV, $E_e \lsim 2$ TeV    &    $\delta E_e \sim 5\% $ \\
&   &  \multirow{2}{*}{(10 ton)}   &   & $\theta_\mu \lsim 0.025$, $\theta_e \lsim 0.5$ &    $\delta \theta_e \sim 15 $ mrad   \\
 &   &     &  & reconstructed $E_h$  &    $\delta E_h \sim 30\% $   \\
  \bottomrule
\end{tabularx}
\vspace{0.2cm}
\caption{\small For each of the far-forward LHC neutrino experiments considered
   we indicate their neutrino pseudo-rapidity coverage, target material, whether
  they can identify the sign of the outgoing charged lepton,
  the acceptance for the charged lepton and hadronic final state,
  and the expected reconstruction performance.
  %
  We consider separately acceptance and performance for electron and muon
  neutrinos.
  %
  For FLArE, we assume that muons would be measured in the FASER2 spectrometer
  situated downstream in the FPF cavern.
  %
  See the description of each experiment in the text for more details.
  %
  For our projections we assume that FASER$\nu$ and SND@LHC acquire data
  for the Run III period ($\mathcal{L}=150$ fb$^{-1}$), while FASER$\nu$2, AdvSND, and FLArE take data
  for the complete HL-LHC period ($\mathcal{L}=3$ ab$^{-1}$).
  \label{tab:FPF_experiments}
}
\end{table}
%-----------------------------------------------------------------
